---
title: "Naïve Bayes Classification"
author: "Skalin Ivan"
date: ""
output: html_document
---

## Assignment 4.5

### a. Construct the Classifier and Display Probabilities

```{r income-model, message=TRUE, warning=FALSE}
library(e1071)

data <- read.csv("./nbtrain.csv")
train <- data[1:9010, ]
test  <- data[9011:10010, ]

model_income <- naiveBayes(income ~ age + sex + educ, data = train)
model_income
```

### if actions are undependened
### P(Action | Atributes) = mul(P(atr_i | Action)) * P(Action) / P(Atributes)
### as we're predicting enum(Actions) for defined Atribures - P(Atributes) is a const so we can ignore it
### we should also use log to not use a lot of multiplication op + _eps, also this will fix having zero in the mul chain
### so: sum(log(P(atr_i | Action) + _eps)) + log(P(Action))

### b. Score the Model and Compute Misclassification Rates

```{r income-evaluation, message=TRUE, warning=FALSE}
pred_income <- predict(model_income, newdata = test)

confusion_matrix <- table(Predicted = pred_income, Actual = test$income)
print("Confusion Matrix for Income Prediction:")
print(confusion_matrix)

overall_error <- mean(pred_income != test$income)
print(paste("Overall Misclassification Rate:", overall_error))

income_levels <- unique(test$income)
for (level in income_levels) {
  total <- sum(confusion_matrix[, level])
  correct <- confusion_matrix[level, level]
  misclass_rate <- 1 - (correct / total)
  print(paste("Misclassification Rate for", level, ":", misclass_rate))
}
```

## Assignment 4.6: Sex Prediction

### a. Build the Classifier and Calculate Misclassification Rates

```{r sex-model, message=TRUE, warning=FALSE}
model_sex <- naiveBayes(sex ~ age + educ + income, data = train)
model_sex

pred_sex <- predict(model_sex, test)

confusion_matrix_sex <- table(Predicted = pred_sex, Actual = test$sex)
print(confusion_matrix_sex)

overall_error_sex <- mean(pred_sex != test$sex)
print(paste("Overall Misclassification Rate:", overall_error_sex))

sex_levels <- unique(test$sex)
for (sex in sex_levels) {
  total <- sum(confusion_matrix_sex[, sex])
  correct <- confusion_matrix_sex[sex, sex]
  misclass_rate <- 1 - (correct / total)
  print(paste("Misclassification Rate for", sex, ":", misclass_rate))
}
```

### b. Balance the Training Data and Reconstruct the Model

```{r sex-balanced-model, message=TRUE, warning=FALSE}
male_data <- subset(train, sex == "M")
female_data <- subset(train, sex == "F")

set.seed(123)
male_sample <- male_data[sample(nrow(male_data), 3500), ]
female_sample <- female_data[sample(nrow(female_data), 3500), ]

train_balanced <- rbind(male_sample, female_sample)

model_balanced <- naiveBayes(sex ~ age + educ + income, data = train_balanced)
model_balanced
```

### c. Evaluate the Balanced Model on Testing Data

```{r sex-balanced-evaluation, message=TRUE, warning=FALSE}
pred_balanced <- predict(model_balanced, test)

confusion_balanced <- table(Predicted = pred_balanced, Actual = test$sex)
cat("Confusion Matrix for Balanced Model:\n")
print(confusion_balanced)

overall_error_balanced <- mean(pred_balanced != test$sex)
print(paste("Overall Misclassification Rate for Balanced Model:", overall_error_balanced))
```

### d. Repeat the Random Sampling Process Multiple Times

```{r sex-multiple-iterations, message=TRUE, warning=FALSE}
for (i in 1:4) {
  male_sample_i   <- male_data[sample(nrow(male_data), 3500), ]
  female_sample_i <- female_data[sample(nrow(female_data), 3500), ]
  train_balanced_i <- rbind(male_sample_i, female_sample_i)
  
  model_i <- naiveBayes(sex ~ age + educ + income, data = train_balanced_i)
  
  pred_i <- predict(model_i, test)
  error_i <- mean(pred_i != test$sex)
  print(paste("Iteration", i, "Overall Misclassification Rate:", error_i))
}

```